Automatic Matching of Bullet Lands

### 2016-01-05 imaging-paper.Rnw
bullet names

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bullet) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bullet~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  #lof <- processBullets(paths = images[c(5,7)], x = 100)

@@

  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bullet~.) +
    geom_line(aes(x=y, y=l30, group=bullet), data=lofX) +
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bullet, linetype=bullet)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y, resid, group=bullet, colour=bullet, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
qplot(y+min(lof$y), resid, group=bullet, colour=bullet, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,250, by=25)
lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

>>>>

\begin{figure}[hbtp]
\centering
    \begin{subfigure}[t]{\textwidth}\centering
\caption{Loess smooth of signatures  at cross section $x = 100$ (span is 0.03).\label{fig:smooth}}{%
<<smooth, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
lofX$bulletname <- scrubPath(lofX$bullet)
qplot(data=lofX, y, resid, colour=I("grey30"), size=I(0.75), geom="line", group=bulletname) +
  geom_line(aes(y=l30), colour="grey70", size=0.5) +
  facet_grid(bulletname~.) +
#  scale_colour_manual("", values=cols) +
  theme_bw() + ylab("") + 
  theme(legend.position="none")
@
    }
\end{subfigure}
\begin{subfigure}[t]{\textwidth}\centering
\caption{Using a rolling median peaks and valleys are identified for each signature. Peaks and valleys on the signature correspond to striation marks on the bullet's surface. \label{fig:smoothcutb}}{%
<<smoothcut, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3.5, out.width='.65\\textwidth', warning=FALSE>>=
#images <- file.path(datadir, dir(datadir))
  #lof <- processBullets(paths = images[c(5,7)], x = 100)

@@

  peaks$bulletname <- scrubPath(peaks$bullet)
  lofX$bulletname <- scrubPath(lofX$bullet)
  ggplot() + theme_bw() +
    geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-6, ymax=6, 
              data=peaks,  alpha=0.2) +
    geom_vline(aes(xintercept=extrema, colour=factor(type)), 
               data= peaks, alpha=0.7) +
    scale_colour_brewer(palette="Set2") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position="none") + 
    facet_grid(bulletname~.) +
    geom_line(aes(x=y, y=l30, group=bulletname), data=lofX) +
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) 
@
}
\end{subfigure}    
\begin{subfigure}[t]{\textwidth}\centering
    \caption{Rectangles in the back identify a striation mark on one of the bullets.  Matching striation marks are indicated by color filled rectangles and marked by an `o'. Mismatches are filled in grey and  marked by an `x'.   \label{fig:smoothcutd}}{%
<<smoothmatch, echo=FALSE, dependson='twolands100match', fig.width=7, fig.height=2.75, out.width='.65\\textwidth', warning=FALSE>>=
peaks1$lines$bullet <- b12[1]
peaks2$lines$bullet <- b12[2]

@@

ggplot() + 
  geom_rect(aes(xmin = xmin, xmax = xmax, fill=factor(type)), ymin = -6, ymax=6.5,  data = lines, alpha=0.2, show.legend = FALSE) +
  theme(legend.position="bottom") +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match)) +
  ylim(c(-6,6.5)) + theme_bw() +
  geom_line(data=lofX, aes(x=y, y=l30, group=bulletname, linetype=bulletname)) +
  scale_linetype_discrete("") +
  scale_colour_manual("", values=cols) +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey60", 0.5)) +
    theme(legend.position = c(1,1.1), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab(expression(paste("Signatures (in ",mu,"m)", sep=""))) +
  xlab("y")
@
    }
\end{subfigure}
\caption{\label{fig:match}Matching striation marks: smooth (a), identify peaks and valley (b), and match peaks and valleys between signatures (c).}
\end{figure}
%\afterpage{\clearpage}
Figure~\ref{fig:match} gives an overview of the automated matching routine: 
We first identify a stable region for each bullet and pick the signature at the lowest cross section level in this region, because typically, individual characteristics are best expressed at the lower end of the bullet (see supplement~\ref{supp:bulletbottom} for a more detailed discussion).

@@

\subsection{Horizontal alignment}
A single cross section across the surface of the two lands in Figure~\ref{fig:manualmatch-rgl} is shown in Figure~\ref{fig:cross100}.  This figure gives the signatures of bullets 1-5 and 2-1 at a surface cross section of $x = 100$. Striation marks show up in this representations as peaks and valleys.  The individual characteristics are quite prominent and, again, suggest a match between the lands. A horizontal shift of one of the signatures (result shown in (b)) emphasizes the strong similarities. 
%
\begin{figure}[hbtp]
\centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{Raw bullet land signatures.\label{fig:crosscut}}{%
<<crosscuts, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lof$bulletname <- scrubPath(lof$bullet)
qplot(y, resid, group=bulletname, colour=bulletname, data=lof, 
      geom="line") + 
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
    \caption{Aligned signatures.\label{fig:crosscutX}}{%
<<crosscutsX, echo=FALSE, dependson='twolands100', fig.width=7, fig.height=3, out.width='\\textwidth'>>=
lofY$bulletname <- scrubPath(lofY$bullet)
qplot(y+min(lof$y), resid, group=bulletname, colour=bulletname, data=lofY, 
      geom="line") + xlab("y") +
  scale_colour_manual("", values=cols) + theme_bw() +
  theme(legend.position = c(1, 1), legend.justification=c("right", "top"),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylab("Residuals of loess fit")
@
    }
\end{subfigure}
\caption{\label{fig:cross100}Signatures of bullets 1-5 and 2-1 at cross sections of $x = 100\mu m$. A horizontal shift of the values of bullet~1-5 to the right shows the similarity of the striation marks.}
\end{figure}
%
This horizontal shift is based on the cross-correlation between the two signatures: let $f(t)$ and $g(t)$ define the signature values  at $t$, where $t$ are locations between 0~$\mu m$ and about 2500~$\mu m$, 1.5625~$\mu m$ apart.
The cross-correlation between $f$ and $g$ at lag $k$ is then defined as
\[
(f * g) (k) = \sum_t f(t+k) g(t),
\]
with suitably defined limits for the summation. % (because variances of $f$ and $g$ do not change, the above equation is proportional to the correlation).

@@

\begin{figure}[hbtp]
  \centering
<<crosscuts-vary-b31, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100,250, by=25)
lof <- processBullets(read.x3p(paths[37]), name = paths[37], x = crosscuts)
lof$bullet <- paste(scrubPath(lof$bullet), lof$x)
